Accessment of Homelessness in Toronto in regards to COVID-19 and other factors

### 2022-02-07 outputs/paper/paper.Rmd
Further dataset cleaning and creating some graphs and tables

The original dataset contained 180 observations with 19 variables. These 180 observations are comprised of 7 categorical entries per month for 24 months (168 observations), and starting from January 2021, an additional entry was added for peoples identifying as Indigenous for a additional 12 observations (totaling 180).

Here is an extract of the dataset with a couple of variables of interest (Table \@ref(tab:extract)).

```{r extract, echo = FALSE, warning = FALSE, message = FALSE}
monthly_shelter_usage %>%
  select(`date(mmm-yy)`, population_group, actively_homeless, population_group_percentage) %>%
  slice(1:7) %>%
  kable(
    caption = "First ten rows of a dataset",
    col.names = c("Date (Month, Year)", "Population Group", "Actively Homeless", "Percentage of Group Population"),
    digits = 1,
    booktabs = TRUE,
    linesep=""
  )
```

Paragraph or two more about it

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
# This needs to be about the random data that I sampled?
```

Talk way more about it.

# Discussion

@@

@@

## Weaknesses and next steps

@@

>>>>

The original dataset contained 180 observations with 18 variables. These 180 observations are comprised of 7 categorical entries per month for 24 months (168 observations) and starting from January 2021, an additional entry was added per month for peoples identifying as Indigenous for an additional 12 observations (totaling 180). For the data cleaning process, variables that are outside of the scope of the report were removed as well as variables that had no potential usage for analysis. The id variable variable was removed due to no potential usage and the 5 age demographic variables were removed because they were outside of the scope for this report and I felt that the Youth categorical observations serve a similar but better statistical purpose. Lastly, I converted the date variable from an (mmm-yy) character into a date format for easier data manipulation and graph creation. This was done using code referenced from [@citeyarna] in the RStudio Community forum.

Here is an extract of the dataset from January 2020, with a few sample of variables of interest (Table \@ref(tab:extract)).

```{r extract, echo = FALSE, warning = FALSE, message = FALSE}
# Showing an extract of the total dataset as a sample
monthly_shelter_usage %>%
  select(date, population_group, actively_homeless, population_group_percentage) %>%
  slice(1:7) %>%
  kable(
    caption = "Extract from full dataset, January 2020",
    col.names = c("Date", "Population Group", "Actively Homeless", "Percentage of Group Population"),
    digits = 1,
    booktabs = TRUE,
    linesep=""
  )
```

As seen, since the cleaned dataset uses the date format, 2020-01-01 represents January, 2020. Moreover, for example, 2021-08-01 would represents August, 2021.

# Data Accessment and Discussion

@@

```{r graph1, fig.cap="Active number of homeless over time", echo=FALSE, fig.height=4, fig.width=6}
# Creating subset of only total population
population_plot <- monthly_shelter_usage[monthly_shelter_usage$population_group == "All Population",]
# Plotting total usage of homeless shelter over time and overlaying start of covid
population_plot %>%
  ggplot(aes(x = date, y = actively_homeless)) +
  labs(x = "Date", y = "Actively Homeless", title =  "Usage of Homeless shelters dropping due to Covid-19") +
  geom_line() +
  geom_line(aes(x=as.Date("2020-03-01")), linetype="dotted", col = "red") + # Drawing dotted red line for covid
  theme_minimal()
```

@@

## Areas of further exploration

@@

Articles that need to be used and referenced in discussion
https://toronto.citynews.ca/2021/03/11/timeline-a-year-of-pandemic-life/
https://www.facetsjournal.com/doi/10.1139/facets-2021-0004
https://www.cbc.ca/news/canada/toronto/toronto-shelters-in-outbreak-unhoused-people-one-death-covid-19-1.6003201
